---
title: "Data Download Vignette"
format:
  html:
    toc: true
    toc-location: right
    embed-resources: true
    code-background: true
    code-tools: true
    code-fold: false
    code-block-border-left: true
    code-link: true
    other-links:
      - text: GitHub Repo
        icon: github
        href: https://github.com/lilykoff/nhanes_steps_mortality
theme: flatly
execute:
  echo: true
  cache: true
  message: false
  warning: false
editor: source
---

The following code can be used to download minute-level step counts and other data from NHANES 2011-2014 from our [Physionet repository](https://physionet.org/content/minute-level-step-count-nhanes/1.0.0/) 


```{r setup}
#| include: false

library(tidyverse)

```


```{r}
#| echo: true
#| eval: false

if(!dir.exists(here::here("data", "accelerometry", "minute_level"))) {
  dir.create(here::here("data", "accelerometry", "minute_level"),
             recursive = TRUE)
}

download_file = function(name) {
  out_file =
    here::here("data",
               "accelerometry",
               "minute_level",
               paste0("nhanes_1440_", name, ".csv.xz"))
  if (!file.exists(out_file)) {
    url = paste0(
      "https://physionet.org/files/minute-level-step-count-nhanes/1.0.0/nhanes_1440_",
      name,
      ".csv.xz"
    )
    system(paste(
      "wget -q --show-progress -O",
      shQuote(out_file),
      shQuote(url)
    ),
    wait = TRUE)
    
  }
}

file_names = c(
  "AC",
  "PAXFLGSM",
  "PAXPREDM",
  "actisteps",
  "adeptsteps",
  "log10AC",
  "log10PAXMTSM",
  "oaksteps",
  "scrfsteps",
  "scsslsteps",
  "vsrevsteps",
  "vssteps"
)
purrr::walk(.x = file_names, .f = download_file)

```

We can read in one file to get a sense of the structure. 
There are columns for ID, day, weekday, and each minute of the day (`min_1` to `min_1440`). Missingness indicates that either: 

1. data were missing for that minute (i.e. wear hadn't started yet)
2. NHANES gave a wear flag for that minute, or
3. the NHANES prediction for that minute was nonwear

```{r}
#| cache: true


sc_steps = read_csv(
  here::here(
    "data",
    "accelerometry",
    "minute_level",
    "nhanes_1440_scsslsteps.csv.xz"
  )
)

sc_steps %>%
  select(
    SEQN,
    PAXDAYM,
    PAXDAYWM,
    min_1,
    min_100,
    min_200,
    min_300,
    min_400,
    min_600,
    min_800,
    min_1000
  ) %>%
  glimpse()
```
